---
title: "All results for the 'Blood Use in HUS' manuscript"
author: "Esa Turkulainen"
date: today
format: 
  html:
    code-fold: true
self-contained: true
---

```{r}
# Libraries n such
library(dplyr)
library(tidyr)
library(purrr)
library(viridis)
library(ggplot2)
library(ggridges)
library(lubridate)
library(reshape2)
library(gridExtra)
library(grid)
library(patchwork)
library(RColorBrewer)
library(ggpattern)
library(boot)

set.seed(123)
```

# Load data

```{r}
master <- readRDS("./results/data/master_blood_use.rds")
```

# Remove outlier

```{r}
master <- master %>%
    filter(id_patient != "xxxx_xxxxxxx")
```

# Time series descriptives

## Intraday series

### Intraday urgency (categories) and EALA (conservative vs operative)

```{r}
# urg
pdf_ts_urgcat <- master %>%
    mutate(hour = hour(start_transf),
           date = as_date(start_transf),
           weekday = wday(start_transf, week_start = 1),
           week = week(start_transf),
           monthday = day(start_transf),
           month = month(start_transf),
           quarter = quarter(start_transf)) %>%
    mutate(date_hour = paste(as.character(date), as.character(hour)),
           date_wday = paste(as.character(date), as.character(weekday)),
           date_mday = paste(as.character(date), as.character(monthday))) %>%
    filter(!is.na(plot_urg_cats)) %>%
    group_by(date_hour, plot_urg_cats) %>%
    mutate(dayhour_count = n()) %>%
    group_by(plot_urg_cats, hour) %>%
    summarise(median_count = median(dayhour_count),
              mean_count = mean(dayhour_count)) %>%
    mutate(max_scaled_mean = mean_count / max(mean_count)) %>%
    ungroup()

# urgcat data N
urgcat_N <- master %>%
    filter(!is.na(plot_urg_cats)) %>%
    nrow() * 100 / nrow(master)

# eala
pdf_ts_ealas <- master %>%
    mutate(hour = hour(start_transf),
           date = as_date(start_transf),
           weekday = wday(start_transf, week_start = 1),
           week = week(start_transf),
           monthday = day(start_transf),
           month = month(start_transf),
           quarter = quarter(start_transf)) %>%
    mutate(date_hour = paste(as.character(date), as.character(hour)),
           date_wday = paste(as.character(date), as.character(weekday)),
           date_mday = paste(as.character(date), as.character(monthday))) %>%
    filter(!is.na(conservative_operative)) %>%
    group_by(date_hour, conservative_operative) %>%
    mutate(dayhour_count = n()) %>%
    group_by(conservative_operative, hour) %>%
    summarise(median_count = median(dayhour_count),
              mean_count = mean(dayhour_count)) %>%
    mutate(max_scaled_mean = mean_count / max(mean_count)) %>%
    ungroup()

# eala_cat N
eala_N <- master %>%
    filter(!is.na(conservative_operative)) %>%
    nrow() * 100 / nrow(master)
```

#### Figure 2 | Part A

```{r}
urg_cat_colors = c("Elective" = "#00BFB2", "Urgent" = "#e7298a")
eala_colors <- c("Conservative" = "#342e37", "Operative" = "#e6ab02")

urg_less_cat_plot <- ggplot(data = pdf_ts_urgcat %>% filter(!(plot_urg_cats %in% c("Organ transplant", "Special resource requirement"))), 
                            aes(x = hour, y = max_scaled_mean, color = plot_urg_cats)) +
    geom_line(size = 1.7) +
    theme_minimal() +
    labs(y = "Relative use",
         x = "Hour of day",
         color = "Urgency:") + 
    scale_color_manual(values = urg_cat_colors) + 
    theme(legend.position = "top")

eala_cat_plot <- ggplot(data = pdf_ts_ealas, aes(x = hour, y = max_scaled_mean, color = conservative_operative)) +
    geom_line(size = 1.4) +
    theme_minimal() +
    labs(y = "Relative use",
         x = "Hour of day",
         color = "Specialty:") + 
    scale_color_manual(values = eala_colors) + 
    theme(legend.position = "top")
```

```{r}
# Combine using patchwork
urg_less_cat_plot / eala_cat_plot + plot_layout(axes = "collect") & theme(legend.position = "top")

ggsave("./results/manuscript1_plots/Figure 2 A.png", width = 2400, height = 1500, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/Figure 2 A.svg", width = 2400, height = 1500, units = "px", bg = "white")
```


## Yearly series (52 weeks)

### Urgency & EALA

```{r}
# urg
pdf_ts_year_urgcat <- master %>%
    mutate(hour = hour(start_transf),
           date = as_date(start_transf),
           weekday = wday(start_transf, week_start = 1),
           week = week(start_transf),
           monthday = day(start_transf),
           month = month(start_transf),
           quarter = quarter(start_transf),
           year = year(start_transf)) %>%
    mutate(date_hour = paste(as.character(date), as.character(hour)),
           date_wday = paste(as.character(date), as.character(weekday)),
           date_mday = paste(as.character(date), as.character(monthday))) %>%
    filter(!is.na(plot_urg_cats)) %>%
    group_by(year, week, plot_urg_cats) %>%
    mutate(week_count = n()) %>%
    group_by(plot_urg_cats, week) %>%
    summarise(median_count = median(week_count),
              mean_count = mean(week_count),
              first_week_count = first(week_count),
              second_week_count = last(week_count)) %>%
    mutate(max_scaled_mean = mean_count / max(mean_count)) %>%
    ungroup() %>%
    filter(week != 53)
# eala
pdf_ts_year_eala <- master %>%
    mutate(hour = hour(start_transf),
           date = as_date(start_transf),
           weekday = wday(start_transf, week_start = 1),
           week = week(start_transf),
           monthday = day(start_transf),
           month = month(start_transf),
           quarter = quarter(start_transf),
           year = year(start_transf)) %>%
    mutate(date_hour = paste(as.character(date), as.character(hour)),
           date_wday = paste(as.character(date), as.character(weekday)),
           date_mday = paste(as.character(date), as.character(monthday))) %>%
    filter(!is.na(conservative_operative)) %>%
    group_by(year, week, conservative_operative) %>%
    mutate(week_count = n()) %>%
    group_by(conservative_operative, week) %>%
    summarise(median_count = median(week_count),
              mean_count = mean(week_count)) %>%
    mutate(max_scaled_mean = mean_count / max(mean_count)) %>%
    ungroup() %>%
    filter(week != 53)
```

#### Supplement Figure S2

```{r}
urg_cat_colors = c("Elective" = "#00BFB2", "Urgent" = "#e7298a")
eala_colors <- c("Conservative" = "#342e37", "Operative" = "#e6ab02")

month <- seq(as.Date("2021-01-01"), 
             as.Date("2021-12-01"), 
             by = "1 month")

month_numeric <- lubridate::yday(month) / 365 * 52 + 1
month_label <- lubridate::month(month, label = TRUE)

month_numeric <- lubridate::yday(month) / 365 * 52 + 1
month_label <- lubridate::month(month, label = TRUE)

urg_less_weekly_plot <- ggplot(data = pdf_ts_year_urgcat %>% filter(!(plot_urg_cats %in% c("Organ transplant", "Special resource requirement"))), aes(x = week, y = max_scaled_mean, color = plot_urg_cats)) +
    geom_line(size = 1.4) +
    labs(y = "Relative use",
         x = "Weeks with month labels",
         color = "Urgency:") + 
    scale_color_manual(values = urg_cat_colors) + 
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_x_continuous(breaks = month_numeric, 
                    labels = month_label)

eala_weekly_plot <- ggplot(data = pdf_ts_year_eala, aes(x = week, y = max_scaled_mean, color = conservative_operative)) +
    geom_line(size = 1.4) +
    theme_minimal() +
    labs(y = "Relative use",
         x = "Weeks with month labels",
         color = "Specialty:") + 
    scale_color_manual(values = eala_colors) + 
    theme(legend.position = "bottom") +
    scale_x_continuous(breaks = month_numeric, 
                    labels = month_label)
```

```{r}
# Combine using patchwork
urg_less_weekly_plot / eala_weekly_plot + plot_layout(axes = "collect") & theme(legend.position = "top")

ggsave("./results/manuscript1_plots/S2.png", width = 2400, height = 1500, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/S2.svg", width = 2400, height = 1500, units = "px", bg = "white")
```

# Urgency categories

## Figure 3 | Part B

```{r}
urgs <- master %>%
    filter(!is.na(urg_op)) %>%
    group_by(urg_op) %>%
    mutate(distinct_episodes_per_urg = n_distinct(id_transfep)) %>%
    mutate(urg_n = n()) %>%
    ungroup() %>%
    mutate(urg_pct = urg_n / n()) %>%
    mutate(urg_per_epis = urg_n / distinct_episodes_per_urg) %>%
    mutate(urg_label = case_when(
        urg_op == "Violetti < 2 h" ~ "2 h",
        urg_op == "Punainen < 6 h" ~ "6 h",
        urg_op == "Oranssi < 24 h" ~ "24 h",
        urg_op == "Keltainen < 48 h" ~ "48 h",
        urg_op == "Vihreä < 7 vrk" ~ "7 d",
        urg_op == "Sininen Elinsiirtokirurgia" ~ "Organ transpl.",
        urg_op == "Ruskea erityisresurssia vaativa" ~ "Special req.",
        urg_op == "LP I" ~ "30 d",
        urg_op == "LP II" ~ "90 d",
        urg_op == "LP III" ~ "180 d"
        )
    ) %>%
    select(urg_label, distinct_episodes_per_urg, urg_n, urg_pct, urg_per_epis) %>%
    distinct() %>%
    mutate(urg_pct_epis = urg_per_epis / sum(urg_per_epis)) %>%
    arrange(desc(urg_n))

urg_order <- c("2 h", "6 h", "24 h", "48 h", "7 d", "Organ transpl.", "Special req.", "30 d", "90 d", "180 d")

urg_N <- master %>%
    filter(!is.na(urg_op)) %>%
    nrow() * 100 / nrow(master)

urg_long <- urgs %>%
    pivot_longer(cols = c("urg_pct", "urg_pct_epis"), names_to = "pct_type", values_to = "pct") %>%
    mutate(pct_type = ifelse(pct_type == "urg_pct", "Total", "Transfusion episodes")) %>%
    mutate(pct_type = factor(pct_type, levels = c("Total", "Transfusion episodes"))) %>%
    mutate(urg_label = factor(urg_label, levels = urg_order))

ggplot(urgs, aes(y = urg_pct, x = factor(urg_label, levels = urg_order), fill = urg_per_epis)) +
    geom_bar(position = "dodge", stat = "identity") +
    geom_text(aes(label = format(urg_n, big.mark = ","), x = urg_label, y = urg_pct + 0.025)) +
    scale_fill_gradient(low = "#fdebf4", high = "#E7298A", name = "Average units per episode") +
    theme_minimal() +
    labs(x = "Urgency", y = "Percent of OR transfusions") +
    theme(legend.position = "bottom") +
    guides(fill = guide_colorbar(barwidth = 20))

ggsave("./results/manuscript1_plots/Figure 3 B.png", width = 3000, height = 1400, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/Figure 3 B.svg", width = 3000, height = 1400, units = "px", bg = "white")

# % of elective category OR transf
master %>%
    filter(!is.na(urg_op)) %>%
    mutate(elec_TRUE = urg_op %in% c("LP I", "LP II", "LP III")) %>%
    summarise(elec_rate = sum(elec_TRUE) / n())
# % of u24 urgency in OR transf
master %>%
    filter(!is.na(urg_op)) %>%
    mutate(urg_TRUE = urg_op %in% c("Violetti < 2 h", "Punainen < 6 h", "Oranssi < 24 h")) %>%
    summarise(urg_rate = sum(urg_TRUE) / n())
# % of organt transplant
master %>%
    mutate(org_TRUE = urg_op == "Sininen Elinsiirtokirurgia") %>%
    summarise(org_rate = sum(org_TRUE, na.rm = T) / n())
# % of organt transplant among OR transf
master %>%
    filter(!is.na(urg_op)) %>%
    mutate(org_TRUE = urg_op == "Sininen Elinsiirtokirurgia") %>%
    summarise(org_rate = sum(org_TRUE) / n())
# % of urgent from OR
sum(urgs[c(1, 3, 4, 5, 7, 9, 10), "urg_n"]) / sum(urgs$urg_n)
# massive transfusions
master %>%
    group_by(id_transfep) %>%
    mutate(massive = n() > 9) %>%
    ungroup() %>%
    summarise(massive_rate = sum(massive) / n())

# Results computed over
100 * sum(urgs$urg_n) / nrow(master) #%
# of units
```

# Desciptives table

## Other stats outside (or for) table

```{r}
# N of product types
prods <- master %>%
    filter(!is.na(type_product)) %>%
    select(type_product) # 160 NA
nRBC <- sum(prods$type_product == "RBC")
nPLT <- sum(prods$type_product == "Platelets")
nPLAS <- sum(prods$type_product == "Plasma")

# radiatedRBC
radiatedRBC <- c("E3847V00", "E7673V00", "A0092VA0", "A0092VB0")
nRadiatedRBC <- master %>%
    filter(veri_admin_prod %in% radiatedRBC) %>%
    nrow()

## Length of hospital stay from apotti_osastohoito
shjakso_lengths <- difftime(master$end_trep_inpeptbl, master$start_trep_inpeptbl, units = "days")
ohjakso_lengths <- difftime(master$end_inpep, master$start_inpep, units = "days")
median(shjakso_lengths, na.rm = T) # median for shjakso: 9.77 days
mean(shjakso_lengths, na.rm = T) # mean for shjakso: 17.84 days
median(ohjakso_lengths, na.rm = T) # median for ohjakso: 1.3 days
quantile(ohjakso_lengths, probs = c(0.25), na.rm = T) # Q25 for ohjakso: 0.39 days (9h22min)
quantile(ohjakso_lengths, probs = c(0.75), na.rm = T) # Q75 for ohjakso: 5.12 days
mean(ohjakso_lengths, na.rm = T) # mean for ohjakso: 4.84 days

## Length of hospital stay from apotti_osastohoito WOMEN
master_women <- master[master$sex_patient == "Nainen", ]
ohjakso_lengths_women <- difftime(master_women$end_inpep, master_women$start_inpep, units = "days")
median(ohjakso_lengths_women, na.rm = T) # median for ohjakso: 1.38
quantile(ohjakso_lengths_women, probs = c(0.25), na.rm = T) # Q25 for ohjakso: 0.38 days (9h7min)
quantile(ohjakso_lengths_women, probs = c(0.75), na.rm = T) # Q75 for ohjakso: 4.86 days

## Length of hospital stay from apotti_osastohoito WOMEN
master_men <- master[master$sex_patient == "Mies", ]
ohjakso_lengths_men <- difftime(master_men$end_inpep, master_men$start_inpep, units = "days")
median(ohjakso_lengths_men, na.rm = T) # median for ohjakso: 1.24
quantile(ohjakso_lengths_men, probs = c(0.25), na.rm = T) # Q25 for ohjakso: 0.4 days (9h36min)
quantile(ohjakso_lengths_men, probs = c(0.75), na.rm = T) # Q75 for ohjakso: 5.69 days

# RBC episodes
master %>%
    group_by(id_transfep) %>%
    mutate(nRBC = sum(type_product == "RBC")) %>%
    filter(nRBC > 0) %>%
    mutate(single = n() == 1) %>%
    slice(1) %>%
    ungroup() %>%
    summarise(singles = sum(single), singlerate = sum(single) / n())

# Plat episodes
master %>%
    group_by(id_transfep) %>%
    mutate(nplat = sum(type_product == "Platelets")) %>%
    filter(nplat > 0) %>%
    mutate(single = n() == 1) %>%
    slice(1) %>%
    ungroup() %>%
    summarise(singles = sum(single), singlerate = sum(single) / n())

```

## Table 1

Currently our columns are "by sex".

```{r}
# Initialize table
table1 <- data.frame(descriptive = character(0), 
                     subgroup = character(0),
                     overall = numeric(0), overall_p = numeric(0), o_Q25 = numeric(0), o_Q75 = numeric(0),
                     men = numeric(0), men_p = numeric(0), men_Q25 = numeric(0), men_Q75 = numeric(0),
                     women = numeric(0), women_p = numeric(0), women_Q25 = numeric(0), women_Q75 = numeric(0))

```

## Number of transfusions

```{r}
# Number of transfusions total
n_transf <- nrow(master)

# By sex
n_transf_women <- nrow(master %>% filter(sex_patient == "Nainen"))
n_transf_men <- nrow(master %>% filter(sex_patient == "Mies")) 

table1 <- bind_rows(table1, 
                    data.frame(descriptive = "Number of transfusions",
                               subgroup = NA,
                               overall = n_transf, overall_p = 100.0, o_Q25 = NA, o_Q75 = NA,
                               men = n_transf_men, men_p = 100 * n_transf_men / n_transf, men_Q25 = NA, men_Q75 = NA,
                               women = n_transf_women, women_p = 100 * n_transf_women / n_transf, women_Q25 = NA, women_Q75 = NA))
```

## Number of transfusion episodes

```{r}
# Number of transfusion episodes total
n_epis <- n_distinct(master$id_transfep)

# By sex
n_epis_women <- n_distinct(master[master$sex_patient == "Nainen", ]$id_transfep)
n_epis_men <- n_distinct(master[master$sex_patient == "Mies", ]$id_transfep)

table1 <- bind_rows(table1, 
                    data.frame(descriptive = "Number of transfusion episodes",
                               subgroup = NA,
                               overall = n_epis, overall_p = 100.0, o_Q25 = NA, o_Q75 = NA,
                               men = n_epis_men, men_p = 100 * n_epis_men / n_epis, men_Q25 = NA, men_Q75 = NA,
                               women = n_epis_women, women_p = 100 * n_epis_women / n_epis, women_Q25 = NA, women_Q75 = NA))
```

## Number of people

```{r}
# Number of unique persons total
n_indv <- n_distinct(master$id_patient); print(n_indv)
# By sex
n_indv_women <- length(unique(master[master$sex_patient == "Nainen", ]$id_patient))
n_indv_men <- length(unique(master[master$sex_patient == "Mies", ]$id_patient))

table1 <- bind_rows(table1, 
                    data.frame(descriptive = "Number of patients", 
                               subgroup = NA,
                               overall = n_indv, overall_p = 100.0, o_Q25 = NA, o_Q75 = NA,
                               men = n_indv_men, men_p = 100 * n_indv_men / n_indv, men_Q25 = NA, men_Q75 = NA,
                               women = n_indv_women, women_p = 100 * n_indv_women / n_indv, women_Q25 = NA, women_Q75 = NA))
```

## Age at transfusion

```{r}
# Age at transfusion
to_agetables <- master %>%
    group_by(id_patient, age_group_patient) %>%
    slice(1) %>%
    ungroup()

agegr_women <- to_agetables %>% filter(sex_patient == "Nainen")
agegr_men <- to_agetables %>% filter(sex_patient == "Mies")

# Overall
transf_ages <- table(to_agetables$age_group_patient)
transf_ages_p <- transf_ages / sum(transf_ages)
# By sex
transf_ages_women <- table(agegr_women$age_group_patient)
transf_ages_women_p <- transf_ages_women / sum(transf_ages_women)
transf_ages_men <- table(agegr_men$age_group_patient)
transf_ages_men_p <- transf_ages_men / sum(transf_ages_men)


agegr_df <- data.frame(descriptive = "Age groups",
                       subgroup = names(transf_ages),
                       overall = as.numeric(transf_ages), overall_p = as.numeric(transf_ages_p), o_Q25 = NA, o_Q75 = NA,
                       men = as.numeric(transf_ages_men), men_p = as.numeric(transf_ages_men_p), men_Q25 = NA, men_Q75 = NA,
                       women = as.numeric(transf_ages_women), women_p = as.numeric(transf_ages_women_p), women_Q25 = NA, women_Q75 = NA)

# Median age overall (+ IQR)
md_age <- median(to_agetables$age_patient)
qs_age <- quantile(to_agetables$age_patient, probs = c(0.25, 0.75))
# By sex
# Women
md_age_women <- median(to_agetables[to_agetables$sex_patient == "Nainen", ]$age_patient)
qs_age_women <- quantile(to_agetables[to_agetables$sex_patient == "Nainen", ]$age_patient, probs = c(0.25, 0.75))
# Men
md_age_men <- median(to_agetables[to_agetables$sex_patient == "Mies", ]$age_patient)
qs_age_men <- quantile(to_agetables[to_agetables$sex_patient == "Mies", ]$age_patient, probs = c(0.25, 0.75))

mdage_df <- data.frame(descriptive = "Age at transfusion (median, IQR)",
                       subgroup = NA,
                       overall = md_age, overall_p = NA, o_Q25 = qs_age[1], o_Q75 = qs_age[2],
                       men = md_age_men, men_p = NA, men_Q25 = qs_age_men[1], men_Q75 = qs_age_men[2],
                       women = md_age_men, women_p = NA, women_Q25 = qs_age_women[1], women_Q75 = qs_age_women[2])

table1 <- bind_rows(table1, agegr_df, mdage_df)
```

## Number of transfusions per patients

Grouped by product type

```{r}
# Number of transfusions per person
transfused_units <- master %>%
    group_by(id_patient, sex_patient, type_product) %>%
    summarise(n_transfused = n()) %>%
    filter(!is.na(type_product))

### RBC
rbcs <- transfused_units %>% 
    filter(type_product == "RBC") 
md_rbc <- median(rbcs$n_transfused)
qs_rbc <- quantile(rbcs$n_transfused, probs = c(0.25, 0.75))
## By sex
# Women
rbcs_w <- transfused_units %>% 
    filter(type_product == "RBC" & sex_patient == "Nainen") 
md_rbc_w <- median(rbcs_w$n_transfused)
qs_rbc_w <- quantile(rbcs_w$n_transfused, probs = c(0.25, 0.75))
# Men
rbcs_m <- transfused_units %>% 
    filter(type_product == "RBC" & sex_patient == "Mies") 
md_rbc_m <- median(rbcs_m$n_transfused)
qs_rbc_m <- quantile(rbcs_m$n_transfused, probs = c(0.25, 0.75))

### Platelets
plats <- transfused_units %>% 
    filter(type_product == "Platelets") 
md_plats <- median(plats$n_transfused)
qs_plats <- quantile(plats$n_transfused, probs = c(0.25, 0.75))
## By sex
# Women
plats_w <- transfused_units %>% 
    filter(type_product == "Platelets" & sex_patient == "Nainen") 
md_plats_w <- median(plats_w$n_transfused)
qs_plats_w <- quantile(plats_w$n_transfused, probs = c(0.25, 0.75))
# Men
plats_m <- transfused_units %>% 
    filter(type_product == "Platelets" & sex_patient == "Mies") 
md_plats_m <- median(plats_m$n_transfused)
qs_plats_m <- quantile(plats_m$n_transfused, probs = c(0.25, 0.75))

### Plasma
plasms <- transfused_units %>% 
    filter(type_product == "Plasma") 
md_plasms <- median(plasms$n_transfused)
qs_plasms <- quantile(plasms$n_transfused, probs = c(0.25, 0.75))
## By sex
# Women
plasms_w <- transfused_units %>% 
    filter(type_product == "Plasma" & sex_patient == "Nainen") 
md_plasms_w <- median(plasms_w$n_transfused)
qs_plasms_w <- quantile(plasms_w$n_transfused, probs = c(0.25, 0.75))
# Men
plasms_m <- transfused_units %>% 
    filter(type_product == "Plasma" & sex_patient == "Mies") 
md_plasms_m <- median(plasms_m$n_transfused)
qs_plasms_m <- quantile(plasms_m$n_transfused, probs = c(0.25, 0.75))

num_by_product_df <- data.frame(descriptive = rep("Per patient (median, IQR)", 3),
                                subgroup = c("RBC", "Platelet", "Plasma"),
                                overall = c(md_rbc, md_plats, md_plasms), overall_p = NA, o_Q25 = c(qs_rbc[1], qs_plats[1], qs_plasms[1]), o_Q75 = c(qs_rbc[2], qs_plats[2], qs_plasms[2]),
                                men = c(md_rbc_m, md_plats_m, md_plasms_m), men_p = NA, men_Q25 = c(qs_rbc_m[1], qs_plats_m[1], qs_plasms_m[1]), men_Q75 = c(qs_rbc_m[2], qs_plats_m[2], qs_plasms_m[2]),
                                women = c(md_rbc_w, md_plats_w, md_plasms_w), women_p = NA, women_Q25 = c(qs_rbc_w[1], qs_plats_w[1], qs_plasms_w[1]), women_Q75 = c(qs_rbc_w[2], qs_plats_w[2], qs_plasms_w[2]))

table1 <- bind_rows(table1, num_by_product_df)
```

## Number of transfusions by EALA

```{r}
ealas <- unique(master$label_primary_specialty)
transfused_eala <- master %>%
    mutate(eala_groups = case_when(
        label_primary_specialty %in% grep("Akuutti", ealas, value = T) ~ "Emergency medicine",
        label_primary_specialty == "20G Kirurgia, gastroenterologia" ~ "Surgery, gastroenterology",
        label_primary_specialty %in% grep("orto", ealas, value = T, ignore.case = T) ~ "Orthopedia",
        label_primary_specialty %in% grep("nefro", ealas, value = T, ignore.case = T) ~ "Nephrology",
        label_primary_specialty %in% grep("hema", ealas, value = T, ignore.case = T) ~ "Hematology",
        label_primary_specialty %in% grep("verisuoni", ealas, value = T, ignore.case = T) ~ "Surgery, vascular",
        label_primary_specialty %in% grep("hema", ealas, value = T, ignore.case = T) ~ "Hematology",
        label_primary_specialty %in% grep("gastro", ealas, value = T, ignore.case = T) & label_primary_specialty != "20G Kirurgia, gastroenterologia" ~ "Gastroenterology",
        label_primary_specialty %in% grep("sydänki", ealas, value = T, ignore.case = T) | label_primary_specialty %in% grep("sydän-", ealas, value = T, ignore.case = T) ~ "Surgery, cardiac",
        label_primary_specialty %in% grep("synn", ealas, value = T, ignore.case = T) ~ "Obstetric",
        label_primary_specialty %in% grep("yleissi", ealas, value = T, ignore.case = T) ~ "General internal medicine",
        label_primary_specialty %in% c("10 Sisätaudit", "Sisätaudit") ~ "Internal medicine",
        label_primary_specialty %in% grep("elinsiirtokirurgia", ealas, value = T, ignore.case = T) ~ "Surgery, transplant",
        label_primary_specialty %in% grep("naistentaudit", ealas, value = T, ignore.case = T) ~ "Gynecology",
        label_primary_specialty %in% grep("plastiikka", ealas, value = T, ignore.case = T) ~ "Surgery, plastic",
        label_primary_specialty %in% grep("urolo", ealas, value = T, ignore.case = T) ~ "Urology",
        label_primary_specialty %in% grep("traumat", ealas, value = T, ignore.case = T) ~ "Surgery, traumatology",
        label_primary_specialty %in% grep("kardio", ealas, value = T, ignore.case = T) ~ "Cardiology",
        label_primary_specialty %in% grep("syöpätaudit ja sädehoito", ealas, value = T, ignore.case = T) ~ "Cancers, radiotherapy",
        label_primary_specialty %in% grep("thorax", ealas, value = T, ignore.case = T) ~ "Surgery, thoracic",
        label_primary_specialty %in% grep("neurokir", ealas, value = T, ignore.case = T) ~ "Surgery, neuro",
        label_primary_specialty %in% grep("infektiosair", ealas, value = T, ignore.case = T) ~ "Infectious diseases",
        label_primary_specialty %in% grep("maksa", ealas, value = T, ignore.case = T) ~ "Surgery, liver",
        label_primary_specialty == "Sydänsiirrot" ~ "Heart transplants",
        TRUE ~ "Other"
    )) %>%
    group_by(eala_groups) %>%
    mutate(eala_sum = n()) %>%
    mutate(eala_pct = eala_sum / nrow(master)) %>%
    group_by(sex_patient) %>%
    mutate(sex_n = n()) %>%
    group_by(eala_groups, sex_patient) %>%
    mutate(eala_sex_sum = n()) %>%
    mutate(eala_sex_pct = eala_sex_sum / sex_n) %>%
    select(eala_groups, eala_sum, eala_pct, sex_patient, sex_n, eala_sex_sum, eala_sex_pct) %>%
    ungroup() %>%
    distinct() %>%
    arrange(desc(eala_sum))

eala_table <- data.frame(descriptive = rep("EALAs", 15),
                         subgroup = unique(transfused_eala$eala_groups)[1:15],
                         overall = unique(transfused_eala$eala_sum)[1:15],
                         overall_p = unique(transfused_eala$eala_pct)[1:15],
                         o_Q25 = NA,
                         o_Q75 = NA,
                         men = transfused_eala[transfused_eala$sex_patient == "Mies", ]$eala_sex_sum[1:15],
                         men_p = transfused_eala[transfused_eala$sex_patient == "Mies", ]$eala_sex_pct[1:15],
                         men_Q25 = NA,
                         men_Q75 = NA,
                         women = transfused_eala[transfused_eala$sex_patient == "Nainen", ]$eala_sex_sum[1:15],
                         women_p = transfused_eala[transfused_eala$sex_patient == "Nainen", ]$eala_sex_pct[1:15],
                         women_Q25 = NA,
                         women_Q75 = NA)

table1 <- bind_rows(table1, eala_table)
```

## Number of transfusions by DRG

```{r}
transfused_DRG <- master %>%
    group_by(code_drg) %>%
    mutate(drg_sum = n()) %>%
    group_by(drg, sex_patient) %>%
    mutate(drg_sex_sum = n()) %>%
    group_by(sex_patient) %>%
    mutate(sex_n = n()) %>%
    select(code_drg, drg_sum, sex_patient, drg_sex_sum, sex_n) %>%
    distinct() %>%
    filter(!is.na(code_drg)) %>%
    ungroup() %>%
    mutate(pct = drg_sum / nrow(master),
           pct_sex = drg_sex_sum / sex_n) %>%
    arrange(desc(drg_sum))

# 13.6.2024
# this below is broken, read results from transfused_DRG instead
drg_table <- data.frame(descriptive = rep("DRGs", 15),
                        subgroup = unique(transfused_DRG$code_drg)[1:15],
                        overall = unique(transfused_DRG$drg_sum)[1:15],
                        overall_p = unique(transfused_DRG$pct)[1:15],
                        o_Q25 = NA,
                        o_Q75 = NA,
                        men = transfused_DRG[transfused_DRG$sex_patient == "Mies", ]$drg_sex_sum[1:15],
                        men_p = transfused_DRG[transfused_DRG$sex_patient == "Mies", ]$pct_sex[1:15],
                        men_Q25 = NA,
                        men_Q75 = NA,
                        women = transfused_DRG[transfused_DRG$sex_patient == "Nainen", ]$drg_sex_sum[1:15],
                        women_p = transfused_DRG[transfused_DRG$sex_patient == "Nainen", ]$pct_sex[1:15],
                        women_Q25 = NA,
                        women_Q75 = NA)

# dirty fix: remove count from men in drg category 372 (Vaginal delivery)
# (the count comes from one row below due to the way we gather the drg_table)
drg_table[drg_table$subgroup == "372", c("men", "men_p")] <- 0

table1 <- bind_rows(table1, drg_table)
saveRDS(table1, "./results/blood_use_Table_1.rds")
```

# Multicomponent transfusions

```{r}
multis <- master %>%
    filter(sex_patient %in% c("Mies", "Nainen")) %>%
    mutate(sex = ifelse(sex_patient == "Mies", "Male", "Female")) %>%
    group_by(id_transfep) %>%
    mutate(epis_n = n()) %>% # needed only for coverage calculation later
    mutate(
        num_RBC = sum(type_product == "RBC", na.rm = T),
        num_Plasma = sum(type_product == "Plasma", na.rm = T),
        num_Platelets = sum(type_product == "Platelets", na.rm = T),
        num_NA_type_product = sum(is.na(type_product))
    ) %>%
    mutate(
        episode_components = case_when(
            num_RBC == 0 & num_Plasma == 0 & num_Platelets == 0 ~ NA,
            num_RBC > 0 & num_Plasma == 0 & num_Platelets == 0 ~ "RBC",
            num_RBC == 0 & num_Plasma > 0 & num_Platelets == 0 ~ "FFP",
            num_RBC == 0 & num_Plasma == 0 & num_Platelets > 0 ~ "PLT",
            num_RBC > 0 & num_Plasma > 0 & num_Platelets == 0 ~ "RBC + FFP",
            num_RBC > 0 & num_Platelets > 0 & num_Plasma == 0 ~ "RBC + PLT",
            num_Plasma > 0 & num_Platelets > 0 & num_RBC == 0 ~ "FFP + PLT",
            num_Plasma > 0 & num_Platelets > 0 & num_RBC > 0 ~ "RBC + FFP + PLT",
            num_NA_type_product > 0 & (num_RBC == 0 | num_Plasma == 0 | num_Platelets == 0) ~ NA
        )
    ) %>%
    slice(1) %>%
    select(sex, id_transfep, episode_components, epis_n) %>%
    ungroup()

plotmultis <- multis %>% # plot w/o NA
    filter(!is.na(episode_components))

multis_N <- 100 * sum(plotmultis$epis_n) / nrow(master)

labelorder <- plotmultis %>%
    group_by(episode_components) %>%
    mutate(epis_comp_n = n()) %>%
    ungroup() %>%
    select(episode_components, epis_comp_n) %>%
    distinct() %>%
    arrange(desc(epis_comp_n))

# Results computed over
100 * sum(multis$epis_n) / nrow(master) #%
# of units
```

## Figure 2 A

Single vs. Multicomponent | yes or no %

```{r}
multis_yesno <- multis %>%
    mutate(single_or_multi = ifelse(episode_components %in% c("RBC", "FFP", "PLT"), "Single component", "Multicomponent"))

ggplot(
    multis_yesno, 
    aes(
        x = factor(single_or_multi, levels = c("Single component", "Multicomponent")), 
        fill = factor(sex, levels = c("Male", "Female")))
    ) +
    geom_bar(
        aes(y = 100 * (..count..)/sum(..count..)),
        position = "dodge"
    ) +
    scale_fill_manual(
        values = gray(c(0.3, 0.7)),
        name = "Sex"
    ) +
    labs(
        x = "", 
        y = "Percent of transfusion episodes"
    ) +
    theme_minimal() +
    theme(
        text = element_text(size = 13),
        legend.position = "bottom"
    )

ggsave(
    "./results/manuscript1_plots/Figure 2 A.png", 
    width = 2700, 
    height = 2100, 
    units = "px", 
    bg = "white"
    )
ggsave(
    "./results/manuscript1_plots/svg/Figure 2 A.svg", 
    width = 2700, 
    height = 2100, 
    units = "px", 
    bg = "white"
    )
```

## Figure 2 B

Single component distirbution | RBC Platelet Plasma %

```{r}
singles <- multis %>%
    filter(episode_components %in% c("RBC", "FFP", "PLT")) %>%
    mutate(episode_components = case_when(
                                    episode_components == "FFP" ~ "Plasma",
                                    episode_components == "PLT" ~ "Platelets",
                                    TRUE ~ "RBC"
    ))

ggplot(
    singles, 
    aes(
        x = factor(episode_components, levels = c("RBC", "Platelets", "Plasma")), 
        fill = factor(sex, levels = c("Male", "Female")))
    ) +
    geom_bar(
        aes(y = 100 * (..count..)/sum(..count..)),
        position = "dodge"
    ) +
    scale_fill_manual(
        values = gray(c(0.3, 0.7)),
        name = "Sex"
    ) +
    labs(
        x = "", 
        y = "Percent of single component transfusion episodes"
    ) +
    theme_minimal() +
    theme(
        text = element_text(size = 13),
        legend.position = "bottom"
    )

ggsave(
    "./results/manuscript1_plots/Figure 2 B.png", 
    width = 2700, 
    height = 2100, 
    units = "px", 
    bg = "white"
    )
ggsave(
    "./results/manuscript1_plots/svg/Figure 2 B.svg", 
    width = 2700, 
    height = 2100, 
    units = "px", 
    bg = "white"
    )
```

## Figure 2 C

Multicomponent distirbution | %

```{r}
only_multis <- multis %>%
    filter(!(episode_components %in% c("RBC", "FFP", "PLT")) & !is.na(episode_components)) %>%
    mutate(episode_components = case_when(
                                    episode_components == "RBC + FFP + PLT" ~ "RBC + Platelets + Plasma",
                                    episode_components == "RBC + PLT" ~ "RBC + Platelets",
                                    episode_components == "RBC + FFP" ~ "RBC + Plasma",
                                    episode_components == "FFP + PLT" ~ "Platelets + Plasma"
    ))

ggplot(
    only_multis, 
    aes(
        x = factor(episode_components, levels = c("RBC + Platelets", "RBC + Plasma", "RBC + Platelets + Plasma", "Platelets + Plasma")), 
        fill = factor(sex, levels = c("Male", "Female")))
    ) +
    geom_bar(
        aes(y = 100 * (..count..)/sum(..count..)),
        position = "dodge"
    ) +
    scale_fill_manual(
        values = gray(c(0.3, 0.7)),
        name = "Sex"
    ) +
    labs(
        x = "", 
        y = "Percent of multicomponent transfusion episodes"
    ) +
    theme_minimal() +
    theme(
        text = element_text(size = 13),
        legend.position = "bottom"
    )

ggsave(
    "./results/manuscript1_plots/Figure 2 C.png", 
    width = 2700, 
    height = 2100, 
    units = "px", 
    bg = "white"
    )
ggsave(
    "./results/manuscript1_plots/svg/Figure 2 C.svg", 
    width = 2700, 
    height = 2100, 
    units = "px", 
    bg = "white"
    )
```

# Transfusions by ICD-10 Category

## Figure 4

```{r}
pdf_icd2 <- master %>%
    filter(!is.na(plot_label_icd_level_2_label) & !is.na(type_product) & sex_patient != "Ei tiedossa/Määriteltävissä") %>%
    group_by(plot_label_icd_level_2_label, sex_patient, type_product) %>%
    summarise(icd_count = n()) %>%
    group_by(plot_label_icd_level_2_label) %>%
    mutate(sum = sum(icd_count)) %>%
    ungroup() %>%
    arrange(desc(sum)) %>%
    mutate(sex_patient = ifelse(sex_patient == "Mies", "Male", "Female"))

icd_N <- master %>%
    filter(!is.na(plot_label_icd_level_2_label) & !is.na(type_product) & sex_patient != "Ei tiedossa/Määriteltävissä") %>%
    nrow() * 100 / nrow(master)

icd2_order <- pdf_icd2 %>%
    distinct(plot_label_icd_level_2_label)

pdf_icd2_topn <- pdf_icd2 %>%
    filter(plot_label_icd_level_2_label %in% icd2_order$plot_label_icd_level_2_label[1:15])

discrete_y <- c(rev(icd2_order$plot_label_icd_level_2_label[1:15]))
# Break long labels
discrete_y_split <- splitAtCentralWhitespace(discrete_y, 33)

pdf_icd2_topn_py <- pdf_icd2_topn %>%
  mutate(x_transformed = ifelse(sex_patient == "Female", 100 * icd_count / nrow(master) * -1, 100 * icd_count / nrow(master)))

# Create a dummy dataframe for the geom_blank
# We need the geom_blank to force free_x scale to be equal
# for both faceted panels
dummy_df <- data.frame(
  plot_label_icd_level_2_label = unique(pdf_icd2_topn$plot_label_icd_level_2_label),
  x_transformed = -7.4,
  type_product = NA,
  sex_patient = "Female"
)

ggplot(pdf_icd2_topn_py,
    aes(
        y = factor(plot_label_icd_level_2_label, levels = rev(icd2_order$plot_label_icd_level_2_label)), 
        x = x_transformed, 
        fill = type_product
    )
    ) +
    geom_bar(stat = "identity") +
    geom_blank(data = dummy_df, aes(x = x_transformed)) +
    # figure out an annotation for faceted plots
    scale_y_discrete(labels = discrete_y_split) +
    scale_fill_manual(values = gray(c(0.3, 0.5, 0.7)), breaks = c("RBC", "Platelets", "Plasma")) +
    facet_wrap(vars(sex_patient), scales = "free_x") +
    theme_minimal() +
    labs(x = "Percent of all transfused units", y = "ICD-10 Category", fill = "Product type") +
    theme(legend.position = "bottom", text = element_text(size = 13), panel.spacing = unit(0, "lines")) +
    scale_x_continuous(labels = abs)


ggsave("./results/manuscript1_plots/Figure 4.png", width = 3000, height = 3000, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/Figure 4.svg", width = 3000, height = 3000, units = "px", bg = "white")

# Share of neoplasms, anemias and cardiovascular diseases
pdf_icd2_topn_py %>%
    filter(plot_label_icd_level_2_label %in% c(
        "C00-97 | Malignant neoplasms",
        "D60-64 | Aplastic and other anaemias",
        "I70-79 | Diseases of arteries, arterioles and capillaries",
        "D37-48 | Neoplasms of uncertain or unknown behaviour",
        "I30-52 | Other forms of heart disease",
        "I20-25 | Ischaemic heart diseases",
        "D50-53 | Nutritional anaemias"
    )) %>%
    distinct(plot_label_icd_level_2_label, sum) %>%
    summarise(pct_tot = sum(sum) / nrow(master))

# Results computed over
100 * sum(pdf_icd2$icd_count) / nrow(master) #%
# of units

# Figure represents
sum(abs(pdf_icd2_topn_py$x_transformed))
```

# Transfusion relation to operation

## Figure 5

```{r}
pdf_liittyy <- master %>%
    mutate(liittyy_extra = if_else(associated_w_or_event != "No", associated_w_or_event, conservative_operative)) %>%
    filter(!is.na(liittyy_extra) & !is.na(type_product)) %>%
    group_by(liittyy_extra, type_product) %>%
    summarise(amount = n()) %>%
    mutate(liittyy_binary = if_else(liittyy_extra %in% c("Conservative", "Operative"), "No", "Yes")) %>%
    ungroup() %>%
    mutate(pct = 100 * amount / nrow(master))

# Summaries in text
# Not-operation-related:
pdf_liittyy %>%
    filter(liittyy_extra %in% c("Conservative", "Operative")) %>%
    summarise(non_pct = sum(pct)) # 52.0%
# Conservative
pdf_liittyy %>%
    filter(liittyy_extra == "Conservative") %>%
    summarise(cons_pct = sum(pct)) # 43.3%
# Conservative out of non-surgical
pdf_liittyy %>%
    filter(liittyy_binary == "No") %>%
    mutate(total = sum(amount)) %>%
    group_by(liittyy_extra) %>%
    summarise(rate = sum(amount) / total) %>%
    distinct()
# Non-surgical operative out of all operative
# assuming all surgical are operative
pdf_liittyy %>%
    filter(liittyy_extra != "Conservative") %>%
    mutate(total = sum(amount)) %>%
    group_by(liittyy_binary) %>%
    summarise(rate = sum(amount) / total)

# When related to operation, preop
pdf_liittyy %>%
    filter(liittyy_extra %in% c("Preop", "Intraop", "Postop")) %>%
    mutate(liittyy_sum = sum(amount)) %>%
    group_by(liittyy_extra) %>%
    mutate(liittyy_pct = 100 * amount / liittyy_sum) %>%
    mutate(liittyy_pct_prodsum = sum(liittyy_pct)) %>%
    select(liittyy_extra, liittyy_pct, liittyy_pct_prodsum)
# 5.0% pre
# 21.4% intra
# 73.6% post

ggplot(pdf_liittyy, aes(x = factor(liittyy_extra, levels = c("Conservative", "Operative", 
                                                             "Preop", "Intraop", "Postop")), 
                        y = pct, 
                        fill = type_product, 
                        pattern = liittyy_binary)) +
    geom_bar_pattern(stat="identity", 
                     pattern_fill = "white",
                     pattern_color = "white",
                     pattern_spacing = 0.03,
                     pattern_frequency = 3, 
                     pattern_angle = 45) +
    scale_pattern_manual(values=c('stripe', 'none')) +
    scale_fill_manual(values = gray(c(0.3, 0.5, 0.7)), breaks = c("RBC", "Platelets", "Plasma")) +
    scale_x_discrete(labels = c("Conservative\nspecialty", "Operative\nspecialty", "Preoperative", "Intraoperative", "Postoperative")) +
    labs(x = "Relation to an operation",
         y = "Percent of transfused units",
         fill = "Product type") +
    theme_minimal() +
    guides(pattern = FALSE,
          fill = guide_legend(override.aes = list(pattern = c("none", "none", "none")))) +
    theme(legend.position = "bottom", text = element_text(size = 13))

ggsave("./results/manuscript1_plots/Figure 5.png", width = 3000, height = 3000, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/Figure 5.svg", width = 3000, height = 3000, units = "px", bg = "white")

# Transfusions associated with a surgical procedure
pdf_liittyy %>%
 filter(liittyy_binary == "Yes") %>%
 summarise(sum = sum(amount)) / sum(pdf_liittyy$amount)

# Results computed over
100 * sum(pdf_liittyy$amount) / nrow(master) #%
# of units
```

# Transfusions by Procedure (Functional-Anatomical Region)

## Figure 6

```{r}
pdf_toim2 <- master %>%
    filter(!is.na(plot_proc_level_2_label) & !is.na(type_product) & sex_patient != "Ei tiedossa/Määriteltävissä" & !(plot_proc_level_2_label %in% grep(" \\| NA", plot_proc_level_2_label, value = T))) %>%
    group_by(plot_proc_level_2_label, sex_patient, type_product) %>%
    summarise(toim_count = n()) %>%
    group_by(plot_proc_level_2_label) %>%
    mutate(sum = sum(toim_count)) %>%
    ungroup() %>%
    arrange(desc(sum)) %>%
    mutate(sex_patient = ifelse(sex_patient == "Mies", "Male", "Female"))

toim2_order <- pdf_toim2 %>%
    distinct(plot_proc_level_2_label)

pdf_toim2_topn <- pdf_toim2 %>%
    filter(plot_proc_level_2_label %in% toim2_order$plot_proc_level_2_label[1:15])

discrete_y <- c(rev(toim2_order$plot_proc_level_2_label[1:15]))
# Break long labels
discrete_y_split <- splitAtCentralWhitespace(discrete_y, 33)

pdf_toim2_topn_py <- pdf_toim2_topn %>%
  mutate(x_transformed = ifelse(sex_patient == "Female", 100 * toim_count / nrow(master) * -1, 100 * toim_count / nrow(master)))

# Create a dummy dataframe for the geom_blank
# We need the geom_blank to force free_x scale to be equal
# for both faceted panels
dummy_df <- data.frame(
  plot_proc_level_2_label = unique(pdf_toim2_topn$plot_proc_level_2_label),
  x_transformed = 3.363,
  type_product = NA,
  sex_patient = "Male"
)

ggplot(pdf_toim2_topn_py, aes(y = factor(plot_proc_level_2_label, levels = rev(toim2_order$plot_proc_level_2_label)), x = x_transformed, fill = factor(type_product, levels = c("Plasma", "Platelets", "RBC")))) +
    geom_bar(stat = "identity") +
    geom_blank(data = dummy_df, aes(x = x_transformed)) +
    # figure out an annotation for faceted plots
    scale_y_discrete(labels = discrete_y_split) +
    scale_fill_manual(values = gray(c(0.3, 0.5, 0.7)), breaks = c("RBC", "Platelets", "Plasma")) +
    facet_wrap(vars(sex_patient), scales = "free_x") +
    theme_minimal() +
    labs(x = "Percent of transfused units", y = "Procedures by functional-anatomical region (NCSP)", fill = "Product type") +
    theme(legend.position = "bottom", text = element_text(size = 13), panel.spacing = unit(0, "lines")) +
    scale_x_continuous(labels = abs)

ggsave("./results/manuscript1_plots/Figure 6.png", width = 3000, height = 3000, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/Figure 6.svg", width = 3000, height = 3000, units = "px", bg = "white")

# Share of NF
master %>%
    mutate(NF = plot_proc_level_2_label == "NF | Hip joint and thigh") %>%
    group_by(sex_patient) %>%
    summarise(NFsum = sum(NF, na.rm = T)) %>%
    mutate(NFrate = NFsum / nrow(master))
# Share of NF + MB + JJ
master %>%
    mutate(
        NFMBJJ = plot_proc_level_2_label %in% c(
            "NF | Hip joint and thigh",
            "MB | Procedures after delivery or abortion",
            "JJ | Liver"
            )
        ) %>%
    group_by(sex_patient) %>%
    summarise(NFMBJJsum = sum(NFMBJJ, na.rm = T)) %>%
    mutate(NFMBJJrate = NFMBJJsum / nrow(master))

# Urgent v. Elective NF
master %>%
    filter(plot_proc_level_2_label == "NF | Hip joint and thigh") %>%
    mutate(urgNF = !(urg_op %in% c("LP I", "LP II", "LP III"))) %>%
    summarise(urgNFsum = sum(urgNF, na.rm = T), urgNFrate = sum(urgNF, na.rm = T) / n())
# Urgent v. Elective NF procedure entities
master %>%
    filter(plot_proc_level_2_label == "NF | Hip joint and thigh") %>%
    group_by(id_proc_set_proctbl) %>% # id for "set of procedures"
    slice(1) %>%
    mutate(urgNF = !(urg_op %in% c("LP I", "LP II", "LP III"))) %>%
    ungroup() %>%
    summarise(urgNFsum = sum(urgNF, na.rm = T), urgNFrate = sum(urgNF, na.rm = T) / n())

# Results computed over
100 * sum(pdf_toim2$toim_count) / nrow(master) #%
# of units

# Figure represents
sum(abs(pdf_toim2_topn_py$x_transformed))
```

# Heatmap of Presentation versus Location

## Figure 3 | Part C

```{r}
presentations_new <- master %>%
    mutate(plot_label_presentation = case_when(
        primary_presentation_method %in% c("Erikoisalasiirto samassa sairaalassa", "Sairaalasiirto") ~ "Hospital or ward transfer",
        primary_presentation_method == "Siirto ajanvarausklinikalta osastohoitoon" & label_where_from_trep == "Sairaala" & label_followup_trep == "Sairaala" ~ "Hospital or ward transfer",
        primary_presentation_method == "Siirto ajanvarausklinikalta osastohoitoon" & !(label_where_from_trep %in% c("Sairaala", "Terveyskeskus", grep("laitos", unique(master$label_where_from_trep), value = T))) & !(label_followup_trep %in% c("Sairaala", "Terveyskeskus", grep("laitos", unique(master$label_followup_trep), value = T))) ~ "Scheduled treatment",
        primary_presentation_method == "Hoidonvaraus" ~ "Scheduled treatment",
        primary_presentation_method == "Päivystys" ~ "Emergency or urgent admission",
        TRUE ~ "Other"
    )) %>%
    mutate(plot_label_place = if_else(primary_location_category %in% c("Other", "Radiology unit", "Home care unit", "Cardiac care unit"), "Other", primary_location_category))

pres_levels <- rev(c("Emergency or urgent admission", "Scheduled treatment", "Hospital or ward transfer", "Other"))
place_levels <- rev(c("Hospital ward", "Operating rooms", "Emergency department", "Intensive care unit", "Outpatient unit",
                      "Dialysis unit", "Endoscopy unit", "Obstetric unit", "Other"))


longtable <- as.data.frame(table(presentations_new$plot_label_place, presentations_new$plot_label_presentation)) # long form table
longtable$Var2 <- factor(longtable$Var2, levels = pres_levels)
longtable$Var1 <- factor(longtable$Var1, levels = place_levels)

# Sum by Var1
sums_var1 <- longtable %>%
  group_by(Var1) %>%
  summarise(Sum = sum(Freq))

# Sum by Var2
sums_var2 <- longtable %>%
  group_by(Var2) %>%
  summarise(Sum = sum(Freq))

# Emergency admission pct
sums_var2[4, "Sum"] / nrow(master) # 61.2 %
sums_var2[4, "Sum"] / sum(sums_var2$Sum) # 61.5 %

# Heatmap coverage
sum(longtable$Freq) / nrow(master)
```

```{r}
# Heatmap plot with no legend or axis titles
heatmap_plot <- ggplot(longtable, aes(x = Var1, y = Var2, fill = Freq)) +
  geom_tile() +
  theme_minimal() +
  scale_fill_gradient(low = "#fdebf4", high = "#E7298A") +
  theme(legend.position = "none", 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank()) +
  geom_text(aes(label = Freq))

# Bar plot for Var1 with no y-axis elements and flipped coordinates
bar_plot_var1 <- ggplot(sums_var1, aes(x = Var1, y = Sum, fill = Sum)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_fill_gradient(low = "#fdebf4", high = "#E7298A") +
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none")

# Bar plot for Var2 with no x-axis elements
bar_plot_var2 <- ggplot(sums_var2, aes(x = Var2, y = Sum, fill = Sum)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_flip() +
  scale_fill_gradient(low = "#fdebf4", high = "#E7298A") +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

# design
design = '
###
A##
BC#
'

combinationplot <- wrap_plots(A = bar_plot_var1, B = heatmap_plot, C = bar_plot_var2, design = design, widths = c(4, 1, 0.1), heights = c(0.1, 1, 2))

combinationplot
  
ggsave("./results/manuscript1_plots/Figure 3 C.png", bg = "white", width = 3000, height = 2200, units = "px")
ggsave("./results/manuscript1_plots/svg/Figure 3 C.svg", width = 3000, height = 1500, units = "px", bg = "white")
```

# Pre and post transfusion Hb 

## Figure 7

```{r}
hbs <- master %>%
    filter(type_product == "RBC") %>%
    group_by(id_transfep) %>%
    arrange(time_hb_pre) %>%
    mutate(hb_tdiff = difftime(time_hb_post_transfep, first(time_hb_pre), units = "hours")) %>%
    mutate(epis_n = n()) %>%
    slice(1) %>%
    ungroup() %>%
    filter(hb_tdiff <= 24) %>%
    select(sex_group_patient, hb_pre, hb_post_transfep, epis_n) %>%
    filter(!is.na(sex_group_patient)) %>%
    filter(!is.na(hb_pre) & !is.na(hb_post_transfep)) %>%
    mutate(sex_group_patient = case_when(sex_group_patient == "mies" ~ "Male",
                                  sex_group_patient == "nainen_postmeno" ~ "Female | Postmenopausal",
                                  sex_group_patient == "nainen_premeno" ~ "Female | Premenopausal")) %>%
    mutate(sex_group_patient = factor(sex_group_patient, levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male"))) %>%
    mutate(hb_delta = (hb_post_transfep - hb_pre) / epis_n) %>%
    group_by(sex_group_patient) %>%
    pivot_longer(
        cols = c(hb_pre, hb_post_transfep), 
        names_to = "variable", 
        values_to = "Hb"
    ) %>%
    mutate(variable = ifelse(variable == "hb_pre", "pre", "post")) %>%
    ungroup()

deltas <- hbs %>%
    select(hb_delta, sex_group_patient) %>%
    filter(row_number() %% 2 == 1)

# Produce boot objects for pre-post-delta by sex_group
## Male
pre_male_bobj <- boot(
    data = hbs[hbs$variable == "pre" & hbs$sex_group_patient == "Male", ]$Hb,
    statistic = md_f,
    R = 10000
)
post_male_bobj <- boot(
    data = hbs[hbs$variable == "post" & hbs$sex_group_patient == "Male", ]$Hb,
    statistic = md_f,
    R = 10000
)
delta_male_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Male", ]$hb_delta,
    statistic = md_f,
    R = 10000
)
## Female | Pre
pre_fempre_bobj <- boot(
    data = hbs[hbs$variable == "pre" & hbs$sex_group_patient == "Female | Premenopausal", ]$Hb,
    statistic = md_f,
    R = 10000
)
post_fempre_bobj <- boot(
    data = hbs[hbs$variable == "post" & hbs$sex_group_patient == "Female | Premenopausal", ]$Hb,
    statistic = md_f,
    R = 10000
)
delta_fempre_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Female | Premenopausal", ]$hb_delta,
    statistic = md_f,
    R = 10000
)
## Female | Post
pre_fempost_bobj <- boot(
    data = hbs[hbs$variable == "pre" & hbs$sex_group_patient == "Female | Postmenopausal", ]$Hb,
    statistic = md_f,
    R = 10000
)
post_fempost_bobj <- boot(
    data = hbs[hbs$variable == "post" & hbs$sex_group_patient == "Female | Postmenopausal", ]$Hb,
    statistic = md_f,
    R = 10000
)
delta_fempost_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Female | Postmenopausal", ]$hb_delta,
    statistic = md_f,
    R = 10000
)

# Produce CIs for bootobjs
## Male
pre_ci_male <- boot.ci(pre_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_male <- boot.ci(post_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_male <- boot.ci(delta_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
## Female | Pre
pre_ci_fempre <- boot.ci(pre_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_fempre <- boot.ci(post_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_fempre <- boot.ci(delta_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
## Female | Post
pre_ci_fempost <- boot.ci(pre_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_fempost <- boot.ci(post_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_fempost <- boot.ci(delta_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca

hbmedians <- data.frame(
    md = c(
        pre_fempre_bobj$t0, 
        post_fempre_bobj$t0,
        delta_fempre_bobj$t0,
        pre_fempost_bobj$t0, 
        post_fempost_bobj$t0,
        delta_fempost_bobj$t0,
        pre_male_bobj$t0, 
        post_male_bobj$t0,
        delta_male_bobj$t0
        ),
    variable = c(rep(c("pre", "post", "pre"), 3)),
    sex_group_patient = factor(c(
        "Female | Premenopausal", "Female | Premenopausal", "Female | Premenopausal",
        "Female | Postmenopausal", "Female | Postmenopausal", "Female | Postmenopausal",
        "Male", "Male", "Male"
        ), levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male")),
    lo = c(
        pre_ci_fempre$percent[4],
        post_ci_fempre$percent[4],        
        delta_ci_fempre$percent[4],
        pre_ci_fempost$percent[4],
        post_ci_fempost$percent[4],        
        delta_ci_fempost$percent[4],
        pre_ci_male$percent[4],
        post_ci_male$percent[4],        
        delta_ci_male$percent[4]
        ),
    hi = c(
        pre_ci_fempre$percent[5],
        post_ci_fempre$percent[5],        
        delta_ci_fempre$percent[5],
        pre_ci_fempost$percent[5],
        post_ci_fempost$percent[5],        
        delta_ci_fempost$percent[5],
        pre_ci_male$percent[5],
        post_ci_male$percent[5],        
        delta_ci_male$percent[5]
        )
)

hbmedians$lo <- round(hbmedians$lo, 1)
hbmedians$hi <- round(hbmedians$hi, 1)

premd <- hbmedians[c(1, 4, 7), ]
postmd <- hbmedians[c(2, 5, 8), ]
deltamd <- hbmedians[c(3, 6, 9), ]

# Ns for plot
hbs_N <- hbs %>%
    group_by(sex_group_patient) %>%
    mutate(N = n() / 2) %>%
    filter(variable == "pre") %>%
    select(sex_group_patient, variable, N) %>%
    distinct()

# df for "Medians [99% CI]" title lol
titles <- data.frame(
    lab = rep("Medians [99% CI]", 3),
    sex_group_patient = factor(c("Female | Premenopausal", "Female | Postmenopausal", "Male"), levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male")),
    variable = rep("pre", 3)
    )

ggplot(hbs, aes(x = Hb, fill = factor(variable, levels = c("pre", "post")))) +
    geom_density() +
    scale_fill_manual(values = gray(c(0.7, 0.3), 0.5), name = "", labels = c("Before transfusion", "After transfusion")) +
    geom_text(data = titles, aes(label = lab, x = 113, y = 0.03), fontface = "bold", size = 3, hjust = 0) +
    geom_text(data = hbs_N, aes(label = paste0("N: ", format(N, big.mark = ",")), x = 30, y = 0.025), fontface = "bold", size = 4, hjust = 0) +
    geom_text(data = premd, 
              aes(label = paste0(
                    "Before transfusion(s): ", md, " [", lo,", ", hi, "]", " g/L"
                    ), 
                  x = 113, 
                  y = 0.025
                 ), 
              size = 3, 
              hjust = 0
             ) +
    geom_text(data = postmd, 
            aes(label = paste0(
                "After transfusion(s): ", md, " [", lo,", ", hi, "]", " g/L"
                ), 
                x = 113, 
                y = 0.020
                ), 
            size = 3, 
            hjust = 0
            ) +
    geom_text(data = deltamd, 
            aes(label = paste0(
                "Δ per transfused unit: +", md, " [", lo,", ", hi, "]", " g/L"
                ), 
                x = 113, 
                y = 0.015
                ), 
            size = 3, 
            hjust = 0
            ) +
    geom_vline(data = premd, aes(xintercept = md), linetype = "dotdash") +
    geom_vline(data = postmd, aes(xintercept = md), linetype = "dashed") +
    facet_wrap(vars(sex_group_patient), ncol = 1) +
    theme_minimal() +
    labs(y = "Density") +
    scale_x_continuous(breaks = c(20, 40, 60, 80, 100, 120, 140, 160)) +
    theme(legend.position = "bottom", strip.text = element_text(size = 13), axis.text = element_text(size = 13), axis.title = element_text(size = 13), legend.text = element_text(size = 13))

ggsave("./results/manuscript1_plots/Figure 7 Col1.png", width = 2000, height = 2000, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/Figure 7 Col1.svg", width = 2000, height = 2000, units = "px", bg = "white")

# Pretransf hb
master %>%
    summarise(md_prehb = median(hb_pre, na.rm = T))

# Results computed over
100 * sum(hbs$epis_n) / 2 / nrow(master) #%
# of units
```


# Bleeding

## Figure 7 (2nd column)

```{r}
bleeds <- master %>%
    filter(!is.na(sex_group_patient) & !is.na(weight_patient) & !is.na(height_patient)) %>%
    filter(bloodloss_pre != 0 | bloodloss_intra != 0) %>%
    mutate(blood_volume = case_when( # Nadler's equation
        sex_patient == "Mies" ~ 1000 * (0.3669 * (height_patient/100)^3 + 0.03219 * weight_patient + 0.6041),
        sex_patient == "Nainen" ~ 1000 * (0.34561 * (height_patient/100)^3 + 0.03308 * weight_patient + 0.1833)
        )
    ) %>%
    mutate(sex_group_patient = case_when(sex_group_patient == "mies" ~ "Male",
                                  sex_group_patient == "nainen_postmeno" ~ "Female | Postmenopausal",
                                  sex_group_patient == "nainen_premeno" ~ "Female | Premenopausal")) %>%
    group_by(id_transfep) %>%
    mutate(units_per_episode = n()) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(bloodloss_pct = 100 * (bloodloss_pre + bloodloss_intra) / blood_volume) %>%
    mutate(bloodloss_pct_per_unit = bloodloss_pct / units_per_episode) %>%
    mutate(sex_group_patient = factor(sex_group_patient, levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male"))) %>%
    select(sex_group_patient, units_per_episode, bloodloss_pct, bloodloss_pct_per_unit)

#######################
# This part for Table 1
bleedstable <- master %>%
    filter(!is.na(sex_group_patient)) %>%
    filter(bloodloss_pre != 0 | bloodloss_intra != 0) %>%
    group_by(id_transfep) %>%
    mutate(units_per_episode = n()) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(bloodloss_per_unit = (bloodloss_pre + bloodloss_intra) / units_per_episode) %>%
    select(sex_patient, bloodloss_per_unit) %>%
    group_by(sex_patient) %>%
    mutate(
        md_bl_unit = median(bloodloss_per_unit),
        range25_bl_unit = quantile(bloodloss_per_unit, probs = c(0.25)),
        range75_bl_unit = quantile(bloodloss_per_unit, probs = c(0.75))
    ) %>%
    ungroup() %>%
    mutate(
        overall_median = median(bloodloss_per_unit),
        overall_range25 = quantile(bloodloss_per_unit, probs = c(0.25)),
        overall_range75 = quantile(bloodloss_per_unit, probs = c(0.75))
    ) %>%
    select(-bloodloss_per_unit) %>%
    distinct()
#######################

bl_male <- bleeds %>%
    filter(sex_group_patient == "Male")
bl_prefem <- bleeds %>%
    filter(sex_group_patient == "Female | Premenopausal")
bl_postfem <- bleeds %>%
    filter(sex_group_patient == "Female | Postmenopausal")

# Extra for text:
bl_fem <- rbind(bl_prefem, bl_postfem)
md_bl_fem <- median(bl_fem$bloodloss_pct)

## Bootstrap pct_per_unit medians
bobj_male <- boot(data = bl_male$bloodloss_pct_per_unit, statistic = md_f, R = 10000) # bump to 10k for final results
bci_male <- boot.ci(bobj_male, type = "bca", conf = 0.99)
lo_male <- bci_male$bca[4]
hi_male <- bci_male$bca[5]

bobj_prefem <- boot(data = bl_prefem$bloodloss_pct_per_unit, statistic = md_f, R = 10000)
bci_prefem <- boot.ci(bobj_prefem, type = "bca", conf = 0.99)
lo_prefem <- bci_prefem$bca[4] 
hi_prefem <- bci_prefem$bca[5]

bobj_postfem <- boot(data = bl_postfem$bloodloss_pct_per_unit, statistic = md_f, R = 10000)
bci_postfem <- boot.ci(bobj_postfem, type = "bca", conf = 0.99)
lo_postfem <- bci_postfem$bca[4] 
hi_postfem <- bci_postfem$bca[5]

## Bootstrap pct medians
bobj_male_pct <- boot(data = bl_male$bloodloss_pct, statistic = md_f, R = 10000) # bump to 10k for final results
bci_male_pct <- boot.ci(bobj_male_pct, type = "bca", conf = 0.99) 
lo_male_pct <- bci_male_pct$bca[4] 
hi_male_pct <- bci_male_pct$bca[5]

bobj_prefem_pct <- boot(data = bl_prefem$bloodloss_pct, statistic = md_f, R = 10000)
bci_prefem_pct <- boot.ci(bobj_prefem_pct, type = "bca", conf = 0.99)
lo_prefem_pct <- bci_prefem_pct$bca[4] 
hi_prefem_pct <- bci_prefem_pct$bca[5]

bobj_postfem_pct <- boot(data = bl_postfem$bloodloss_pct, statistic = md_f, R = 10000)
bci_postfem_pct <- boot.ci(bobj_postfem_pct, type = "bca", conf = 0.99)
lo_postfem_pct <- bci_postfem_pct$bca[4] 
hi_postfem_pct <- bci_postfem_pct$bca[5]

bootmedians <- data.frame(
    md_pct_unit = c(
        bobj_prefem$t0, 
        bobj_postfem$t0,
        bobj_male$t0
        ),
    md_pct = c(
        bobj_prefem_pct$t0, 
        bobj_postfem_pct$t0,
        bobj_male_pct$t0
    ),
    sex_group_patient = factor(c(
        "Female | Premenopausal",
        "Female | Postmenopausal",
        "Male"
        )),
    lo_unit = c(
        lo_prefem,
        lo_postfem,        
        lo_male
        ),
    hi_unit = c(
        hi_prefem,
        hi_postfem,
        hi_male
        ),
    lo_pct = c(
        lo_prefem_pct,
        lo_postfem_pct,        
        lo_male_pct
    ),
    hi_pct = c(
        hi_prefem_pct,
        hi_postfem_pct,
        hi_male_pct
    )
)

bleeds_N <- bleeds %>%
    group_by(sex_group_patient) %>%
    summarise(N = n()) %>%
    select(sex_group_patient, N) %>%
    distinct()

ggplot(bleeds, aes(x = bloodloss_pct_per_unit)) +
    geom_density(fill = gray(0.7), alpha = 0.5) +
    facet_wrap(vars(sex_group_patient), ncol = 1) +
    xlim(c(0, 20)) +
    geom_text(data = bootmedians[bootmedians$sex_group_patient != "Female | Premenopausal", ], aes(label = "Medians [99% CI]", fontface = "bold", x = 10, y = 0.10), size = 3, hjust = 0) +
    geom_text(data = bootmedians[bootmedians$sex_group_patient == "Female | Premenopausal", ], aes(label = "Medians [99% CI]", fontface = "bold", x = 0, y = 0.105), size = 3, hjust = 0) +
    geom_text(data = bleeds_N, aes(label = paste("N:", format(N, big.mark = ",", trim = T)), x = 15, y = 0.05), fontface = "bold", size = 4, hjust = 0) +
    geom_text(data = bootmedians[bootmedians$sex_group_patient == "Female | Premenopausal", ], aes(label = paste0(
                    "Loss per episode: ", round(md_pct, 2), " [", round(lo_pct, 2), ", ", round(hi_pct, 2), "]", " %\n", 
                    "Loss per transfused unit: ", round(md_pct_unit, 2), " [", round(lo_unit, 2), ", ", round(hi_unit, 2), "]", " %"
                    ), 
                  x = 0, 
                  y = 0.085
                 ), 
              size = 3, 
              hjust = 0) +
    geom_text(data = bootmedians[bootmedians$sex_group_patient != "Female | Premenopausal", ], 
              aes(label = paste0(
                    "Loss per episode: ", round(md_pct, 2), " [", round(lo_pct, 2), ", ", round(hi_pct, 2), "]", " %\n", 
                    "Loss per transfused unit: ", round(md_pct_unit, 2), " [", round(lo_unit, 2), ", ", round(hi_unit, 2), "]", " %"
                    ), 
                  x = 10, 
                  y = 0.08
                 ), 
              size = 3, 
              hjust = 0
             ) +
    geom_vline(data = bootmedians, aes(xintercept = md_pct_unit), linetype = "dashed") +
    theme_minimal() +
    labs(x = "Loss of blood (% of volume) per transfused unit",
         y = "Density") +
    theme(strip.text = element_text(size = 13), axis.text = element_text(size = 13), axis.title = element_text(size = 13), legend.text = element_text(size = 13))

ggsave("./results/manuscript1_plots/Figure 7 Col2.png", width = 2000, height = 2000, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/Figure 7 Col2.svg", width = 2000, height = 2000, units = "px", bg = "white")

# Results computed over
100 * sum(bleeds$units_per_episode) / nrow(master) #%
# of units
```

# Massive transfusion

Currently only in text

```{r}
massives <- master %>%
    group_by(id_transfep) %>%
    mutate(episode_n = n()) %>%
    ungroup() %>%
    mutate(massive = ifelse(episode_n > 9, TRUE, FALSE)) %>%
    filter(massive) %>%
    group_by(sex_group_patient, type_product) %>%
    mutate(distinct_episodes = n_distinct(id_transfep)) %>%
    mutate(massive_n = n()) %>%
    mutate(massive_per_epi = massive_n / distinct_episodes) %>%
    select(sex_group_patient, distinct_episodes, massive_n, massive_per_epi) %>%
    distinct() %>%
    ungroup() %>%
    mutate(
        massive_pct = 100 * massive_n / sum(massive_n),
        massive_pct_epis = 100 * massive_per_epi / sum(massive_per_epi)
    ) %>%
    arrange(desc(massive_n))
```

# Supplemental

## From first diagnosis to transfusion

```{r}
diags <- readRDS("./results/data/apotti_diagnoosi.rds") %>%
    filter(henkilotunnus %in% unique(master$id_patient)) %>%
    filter(paadiagnoosi) %>%
    rename("id_patient" = "henkilotunnus")
```

```{r}
diags_sel <- diags %>%
    select(id_patient, diagnoosi_koodi, ilmoitettu_alkuhetki) %>%
    group_by(id_patient, diagnoosi_koodi) %>%
    slice_min(order_by = ilmoitettu_alkuhetki)
rm(diags)
pat_ICD <- master %>%
    select(id_patient, code_main_icd_contact, start_transf, label_icd_level_1, label_icd_level_2) %>%
    group_by(id_patient, code_main_icd_contact) %>%
    slice_min(order_by = start_transf)

others <- c("Factors influencing health status and contact with health services",
            "Mental and behavioural disorders",
            "Codes for special purposes")

pat_diags <- pat_ICD %>%
    left_join(diags_sel, by = c("id_patient", "code_main_icd_contact" = "diagnoosi_koodi")) %>%
    mutate(diag_to_transf_hours = as.numeric(difftime(start_transf, ilmoitettu_alkuhetki, units = "hours"))) %>%
    mutate(diag_to_transf_days = diag_to_transf_hours / 24) %>%
    filter(diag_to_transf_hours >= 0) %>%
    group_by(label_icd_level_1) %>%
    #group_by(label_icd_level_2) %>%
    mutate(mean_tdiff_hours = mean(diag_to_transf_hours)) %>%
    mutate(mean_tdiff_days = mean_tdiff_hours / 24) %>%
    mutate(median_tdiff_hours = median(diag_to_transf_hours)) %>%
    mutate(median_tdiff_days = median_tdiff_hours / 24) %>%
    filter(!is.na(label_icd_level_1) & label_icd_level_1 != "Diseases of the ear and mastoid process" & label_icd_level_1 != "Diseases of the eye and adnexa") %>%
    mutate(ICD_plotlabel = case_when(
        label_icd_level_1 == "Diseases of the blood and blood-forming organs and certain disorders involving the immune mechanism" ~ "Diseases of blood & immune mechanism",
        label_icd_level_1 %in% others ~ "Other",
        TRUE ~ label_icd_level_1)
        ) %>%
    group_by(ICD_plotlabel) %>%
    mutate(ICD1_num = n()) %>%
    mutate(label_median_days = median(diag_to_transf_days)) %>%
    mutate(ICD2_num = n()) %>%
    mutate(median_days = median(diag_to_transf_days)) %>%
    ungroup()
rm(pat_ICD)
labelorder <- pat_diags %>%
    distinct(ICD_plotlabel, median_days) %>%
    arrange(desc(median_days))

discrete_y <- c(rev(labelorder$ICD_plotlabel[1:15]))
# Break long labels
discrete_y_split <- splitAtCentralWhitespace(discrete_y, 33)

pat_diags$ICD_plotlabel <- factor(pat_diags$ICD_plotlabel, levels = labelorder$ICD_plotlabel)
```

```{r}
diag_N = pat_diags %>%
    select(ICD_plotlabel, ICD1_num) %>%
    distinct()

ggplot(
    pat_diags %>% filter(diag_to_transf_days > 0), 
    aes(
        x = diag_to_transf_days, 
        y = ICD_plotlabel, 
        fill = log(..x..)
        )
    ) +
    geom_density_ridges_gradient(
        scale = 3, 
        rel_min_height = 0.01, 
        quantile_lines = T, 
        quantiles = 2,
        alpha = 0.4
        ) +
    geom_text(
        data = diag_N, 
        aes(
            x = 4000, 
            y = ICD_plotlabel, 
            label = paste0("N = ", format(ICD1_num, big.mark = ","))), 
            nudge_y = 0.5,
            fontface = "bold"
            ) +
    scale_fill_gradient(low = "#fdebf4", high = "#E7298A") +
    scale_y_discrete(labels = discrete_y_split) +
    theme_minimal() +
    theme(
        legend.position = "none",
        panel.spacing = unit(0.1, "lines"),
        text = element_text(size = 13)
        ) +
    scale_x_continuous(
        breaks = c(1, 7, 30, 365, 3650), 
        limits = c(0.1, 20000), 
        trans = "log"
        ) +
    labs(
        x = "Days",
        y = "ICD-10 Top Level Categories"
        )

ggsave("./results/manuscript1_plots/S3.png", bg = "white", width = 3000, height = 3000, units = "px")
ggsave("./results/manuscript1_plots/svg/S3.svg", bg = "white", width = 3000, height = 3000, units = "px")
```

## Number of transfused units -distributions

```{r}
fill_colors <- c("#e7298a", "#00BFB2", "#4e4553")
color_colors <- c("#c71670", "#008c82", "#342e37")

master_filtered <- master %>% 
    filter(!is.na(type_product))

master_filtered$type_product <- factor(master_filtered$type_product, levels = c("RBC", "Platelets", "Plasma"))
master$type_product <- factor(master$type_product, levels = c("RBC", "Platelets", "Plasma"))

# Episode wise
p_episode <- transfusionEventDistribution_Grouped(
    master,
    variable="id_transfep",
    grouping_variable = "type_product",
    color_var_name = "blood product",
    color_var_name_x = "blood product",
    y_var_name = "transfusion episodes",
    fill_var_name = "transfusion\nepisodes",
    grouping_var_name = "product type",
    fill_colors = fill_colors,
    color_colors = color_colors
)

p_episode <- p_episode + 
    theme(legend.position = "bottom", legend.box="horizontal") +
    guides(fill = guide_legend(nrow = 3, order = 1), color = guide_legend(nrow = 3, order = 2))

p_episode

ggsave("./results/manuscript1_plots/S4_A.png", width = 2300, height = 2500, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/S4_A.svg", width = 2300, height = 2500, units = "px", bg = "white")

# Shjakso wise
p_shj <- transfusionEventDistribution_Grouped(
    master_filtered,
    variable="shjakso_numero",
    grouping_variable = "type_product",
    color_var_name = "blood product",
    color_var_name_x = "blood product",
    y_var_name = "treatment episodes",
    fill_var_name = "treatment\nepisodes",
    grouping_var_name = "product type",
    fill_colors = fill_colors,
    color_colors = color_colors
)
p_shj <- p_shj + 
    theme(legend.position = "bottom", legend.box="horizontal") +
    guides(fill = guide_legend(nrow = 3, order = 1), color = guide_legend(nrow = 3, order = 2))
p_shj

ggsave("./results/manuscript1_plots/S4_B.png", width = 2300, height = 2500, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/S4_B.svg", width = 2300, height = 2500, units = "px", bg = "white")
```

## Trom labs

Code adapted completely from analyses_included.qmd by Esa.

```{r}
Troms <- master %>%
    filter(type_product == "Platelets") %>%
    group_by(id_transfep) %>%
    arrange(time_plt_pre) %>%
    mutate(Trom_tdiff = difftime(time_plt_post_transfep, first(time_plt_pre), units = "hours")) %>%
    mutate(epis_n = n()) %>%
    slice(1) %>%
    ungroup() %>%
    filter(Trom_tdiff <= 24) %>%
    select(sex_group_patient, plt_pre, plt_post_transfep, epis_n) %>%
    filter(!is.na(sex_group_patient)) %>%
    filter(!is.na(plt_pre) & !is.na(plt_post_transfep)) %>%
    mutate(sex_group_patient = case_when(sex_group_patient == "mies" ~ "Male",
                                  sex_group_patient == "nainen_postmeno" ~ "Female | Postmenopausal",
                                  sex_group_patient == "nainen_premeno" ~ "Female | Premenopausal")) %>%
    mutate(sex_group_patient = factor(sex_group_patient, levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male"))) %>%
    mutate(Trom_delta = (plt_post_transfep - plt_pre) / epis_n) %>%
    group_by(sex_group_patient) %>%
    pivot_longer(
        cols = c(plt_pre, plt_post_transfep), 
        names_to = "variable", 
        values_to = "Trom"
    ) %>%
    mutate(variable = ifelse(variable == "plt_pre", "pre", "post")) %>%
    ungroup()

deltas <- Troms %>%
    select(Trom_delta, sex_group_patient) %>%
    filter(row_number() %% 2 == 1)

## This for text:
Troms %>%
    filter(sex_group_patient != "Male") %>%
    filter(variable == "pre") %>%
    summarise(
        md = median(Trom),
        Q25 = quantile(Trom, 0.25),
        Q75 = quantile(Trom, 0.75))
Troms %>%
    filter(sex_group_patient == "Male") %>%
    filter(variable == "pre") %>%
    summarise(
        md = median(Trom),
        Q25 = quantile(Trom, 0.25),
        Q75 = quantile(Trom, 0.75))

# Produce boot objects for pre-post-delta by sex_group
## Male
pre_male_bobj <- boot(
    data = Troms[Troms$variable == "pre" & Troms$sex_group_patient == "Male", ]$Trom,
    statistic = md_f,
    R = 10000
)
post_male_bobj <- boot(
    data = Troms[Troms$variable == "post" & Troms$sex_group_patient == "Male", ]$Trom,
    statistic = md_f,
    R = 10000
)
delta_male_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Male", ]$Trom_delta,
    statistic = md_f,
    R = 10000
)
## Female | Pre
pre_fempre_bobj <- boot(
    data = Troms[Troms$variable == "pre" & Troms$sex_group_patient == "Female | Premenopausal", ]$Trom,
    statistic = md_f,
    R = 10000
)
post_fempre_bobj <- boot(
    data = Troms[Troms$variable == "post" & Troms$sex_group_patient == "Female | Premenopausal", ]$Trom,
    statistic = md_f,
    R = 10000
)
delta_fempre_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Female | Premenopausal", ]$Trom_delta,
    statistic = md_f,
    R = 10000
)
## Female | Post
pre_fempost_bobj <- boot(
    data = Troms[Troms$variable == "pre" & Troms$sex_group_patient == "Female | Postmenopausal", ]$Trom,
    statistic = md_f,
    R = 10000
)
post_fempost_bobj <- boot(
    data = Troms[Troms$variable == "post" & Troms$sex_group_patient == "Female | Postmenopausal", ]$Trom,
    statistic = md_f,
    R = 10000
)
delta_fempost_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Female | Postmenopausal", ]$Trom_delta,
    statistic = md_f,
    R = 10000
)

# Produce CIs for bootobjs
## Male
pre_ci_male <- boot.ci(pre_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_male <- boot.ci(post_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_male <- boot.ci(delta_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
## Female | Pre
pre_ci_fempre <- boot.ci(pre_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_fempre <- boot.ci(post_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_fempre <- boot.ci(delta_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
## Female | Post
pre_ci_fempost <- boot.ci(pre_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_fempost <- boot.ci(post_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_fempost <- boot.ci(delta_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca

Trommedians <- data.frame(
    md = c(
        pre_fempre_bobj$t0, 
        post_fempre_bobj$t0,
        delta_fempre_bobj$t0,
        pre_fempost_bobj$t0, 
        post_fempost_bobj$t0,
        delta_fempost_bobj$t0,
        pre_male_bobj$t0, 
        post_male_bobj$t0,
        delta_male_bobj$t0
        ),
    variable = c(rep(c("pre", "post", "pre"), 3)),
    sex_group_patient = factor(c(
        "Female | Premenopausal", "Female | Premenopausal", "Female | Premenopausal",
        "Female | Postmenopausal", "Female | Postmenopausal", "Female | Postmenopausal",
        "Male", "Male", "Male"
        ), levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male")),
    lo = c(
        pre_ci_fempre$percent[4],
        post_ci_fempre$percent[4],        
        delta_ci_fempre$percent[4],
        pre_ci_fempost$percent[4],
        post_ci_fempost$percent[4],        
        delta_ci_fempost$percent[4],
        pre_ci_male$percent[4],
        post_ci_male$percent[4],        
        delta_ci_male$percent[4]
        ),
    hi = c(
        pre_ci_fempre$percent[5],
        post_ci_fempre$percent[5],        
        delta_ci_fempre$percent[5],
        pre_ci_fempost$percent[5],
        post_ci_fempost$percent[5],        
        delta_ci_fempost$percent[5],
        pre_ci_male$percent[5],
        post_ci_male$percent[5],        
        delta_ci_male$percent[5]
        )
)

Trommedians$lo <- round(Trommedians$lo, 1)
Trommedians$hi <- round(Trommedians$hi, 1)

premd <- Trommedians[c(1, 4, 7), ]
postmd <- Trommedians[c(2, 5, 8), ]
deltamd <- Trommedians[c(3, 6, 9), ]

# 0.975 quantile for x-limits
max_q975 <- Troms %>%
    group_by(sex_group_patient, variable) %>%
    summarise(q975 = quantile(Trom, p=0.975)) %>%
    pull(q975) %>%
    max()

# Ns for plot
Troms_N <- Troms %>%
    filter(Trom <= max_q975) %>%
    group_by(sex_group_patient) %>%
    mutate(N = round(n() / 2)) %>%
    select(sex_group_patient, variable, N) %>%
    distinct()

# df for "Medians [99% CI]" title lol
titles <- data.frame(
    lab = rep("Medians [99% CI]", 3),
    sex_group_patient = factor(c("Female | Premenopausal", "Female | Postmenopausal", "Male"), levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male")),
    variable = rep("pre", 3)
    )

ggplot(Troms, aes(x = Trom, fill = factor(variable, levels = c("pre", "post")))) +
    geom_density() +
    scale_fill_manual(values = gray(c(0.7, 0.3), 0.5), name = "", labels = c("Before transfusion", "After transfusion")) +
    geom_text(data = titles, aes(label = "Medians [99% CI]", fontface = "bold", x = 150, y = 0.015), size = 3, hjust = 0) +
    geom_text(data = Troms_N, aes(label = paste("N:", format(N, big.mark = ",", trim = T)), x = 15, y = 0.012), fontface = "bold", size = 4, hjust = 0) +
    geom_text(data = premd, 
              aes(label = paste0(
                    "Before transfusion(s): ", md, " [", lo,", ", hi, "]", " E9/L"
                    ), 
                  x = 150, 
                  y = 0.012
                 ), 
              size = 3, 
              hjust = 0
             ) +
    geom_text(data = postmd, 
            aes(label = paste0(
                "After transfusion(s): ", md, " [", lo,", ", hi, "]", " E9/L"
                ), 
                x = 150, 
                y = 0.009
                ), 
            size = 3, 
            hjust = 0
            ) +
    geom_text(data = deltamd, 
            aes(label = paste0(
                "Δ per transfused unit: +", md, " [", lo,", ", hi, "]", " E9/L"
                ), 
                x = 150, 
                y = 0.006
                ), 
            size = 3, 
            hjust = 0
            ) +
    geom_vline(data = premd, aes(xintercept = md), linetype = "dotdash") +
    geom_vline(data = postmd, aes(xintercept = md), linetype = "dashed") +
    facet_wrap(vars(sex_group_patient), ncol = 1) +
    theme_minimal() +
    labs(y = "Density", x="Platelet count") +
    theme(legend.position = "bottom", strip.text = element_text(size = 13), axis.text = element_text(size = 13), axis.title = element_text(size = 13), legend.text = element_text(size = 13), axis.title.x = element_text(margin = margin(t = 14)),) +
    scale_x_continuous(breaks = seq(0, max_q975, by=50), limits = c(0, max_q975))

ggsave("./results/manuscript1_plots/S5_A.png", width = 2000, height = 2000, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/S5_A.svg", width = 2000, height = 2000, units = "px", bg = "white")

# Results computed over
100 * sum(Troms$epis_n) / 2 / nrow(master) #%
# of units
```

## INR labs

```{r}
INRs <- master %>%
    filter(type_product == "Plasma") %>%
    group_by(id_transfep) %>%
    arrange(time_inr_pre) %>%
    mutate(INR_tdiff = difftime(time_inr_post_transfep, first(time_inr_pre), units = "hours")) %>%
    mutate(epis_n = n()) %>%
    slice(1) %>%
    ungroup() %>%
    filter(INR_tdiff <= 24) %>%
    select(sex_group_patient, inr_pre, inr_post_transfep, epis_n) %>%
    filter(!is.na(sex_group_patient)) %>%
    filter(!is.na(inr_pre) & !is.na(inr_post_transfep)) %>%
    mutate(sex_group_patient = case_when(sex_group_patient == "mies" ~ "Male",
                                  sex_group_patient == "nainen_postmeno" ~ "Female | Postmenopausal",
                                  sex_group_patient == "nainen_premeno" ~ "Female | Premenopausal")) %>%
    mutate(sex_group_patient = factor(sex_group_patient, levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male"))) %>%
    mutate(INR_delta = (inr_post_transfep - inr_pre) / epis_n) %>%
    group_by(sex_group_patient) %>%
    pivot_longer(
        cols = c(inr_pre, inr_post_transfep), 
        names_to = "variable", 
        values_to = "INR"
    ) %>%
    mutate(variable = ifelse(variable == "inr_pre", "pre", "post")) %>%
    ungroup()

deltas <- INRs %>%
    select(INR_delta, sex_group_patient) %>%
    filter(row_number() %% 2 == 1)

## For text:
INRs %>%
    filter(sex_group_patient != "Male") %>%
    filter(variable == "pre") %>%
    summarise(
        md = median(INR),
        Q25 = quantile(INR, 0.25),
        Q75 = quantile(INR, 0.75)
    )
INRs %>%
    filter(sex_group_patient == "Male") %>%
    filter(variable == "pre") %>%
    summarise(
        md = median(INR),
        Q25 = quantile(INR, 0.25),
        Q75 = quantile(INR, 0.75)
    )
# Produce boot objects for pre-post-delta by sex_group
## Male
pre_male_bobj <- boot(
    data = INRs[INRs$variable == "pre" & INRs$sex_group_patient == "Male", ]$INR,
    statistic = md_f,
    R = 10000
)
post_male_bobj <- boot(
    data = INRs[INRs$variable == "post" & INRs$sex_group_patient == "Male", ]$INR,
    statistic = md_f,
    R = 10000
)
delta_male_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Male", ]$INR_delta,
    statistic = md_f,
    R = 10000
)
## Female | Pre
pre_fempre_bobj <- boot(
    data = INRs[INRs$variable == "pre" & INRs$sex_group_patient == "Female | Premenopausal", ]$INR,
    statistic = md_f,
    R = 10000
)
post_fempre_bobj <- boot(
    data = INRs[INRs$variable == "post" & INRs$sex_group_patient == "Female | Premenopausal", ]$INR,
    statistic = md_f,
    R = 10000
)
delta_fempre_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Female | Premenopausal", ]$INR_delta,
    statistic = md_f,
    R = 10000
)
## Female | Post
pre_fempost_bobj <- boot(
    data = INRs[INRs$variable == "pre" & INRs$sex_group_patient == "Female | Postmenopausal", ]$INR,
    statistic = md_f,
    R = 10000
)
post_fempost_bobj <- boot(
    data = INRs[INRs$variable == "post" & INRs$sex_group_patient == "Female | Postmenopausal", ]$INR,
    statistic = md_f,
    R = 10000
)
delta_fempost_bobj <- boot(
    data = deltas[deltas$sex_group_patient == "Female | Postmenopausal", ]$INR_delta,
    statistic = md_f,
    R = 10000
)

# Produce CIs for bootobjs
## Male
pre_ci_male <- boot.ci(pre_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_male <- boot.ci(post_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_male <- boot.ci(delta_male_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
## Female | Pre
pre_ci_fempre <- boot.ci(pre_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_fempre <- boot.ci(post_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_fempre <- boot.ci(delta_fempre_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
## Female | Post
pre_ci_fempost <- boot.ci(pre_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
post_ci_fempost <- boot.ci(post_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca
delta_ci_fempost <- boot.ci(delta_fempost_bobj, type = "perc", conf = 0.99) # estimates don't vary enough to use bca

INRmedians <- data.frame(
    md = c(
        pre_fempre_bobj$t0, 
        post_fempre_bobj$t0,
        delta_fempre_bobj$t0,
        pre_fempost_bobj$t0, 
        post_fempost_bobj$t0,
        delta_fempost_bobj$t0,
        pre_male_bobj$t0, 
        post_male_bobj$t0,
        delta_male_bobj$t0
        ),
    variable = c(rep(c("pre", "post", "pre"), 3)),
    sex_group_patient = factor(c(
        "Female | Premenopausal", "Female | Premenopausal", "Female | Premenopausal",
        "Female | Postmenopausal", "Female | Postmenopausal", "Female | Postmenopausal",
        "Male", "Male", "Male"
        ), levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male")),
    lo = c(
        pre_ci_fempre$percent[4],
        post_ci_fempre$percent[4],        
        delta_ci_fempre$percent[4],
        pre_ci_fempost$percent[4],
        post_ci_fempost$percent[4],        
        delta_ci_fempost$percent[4],
        pre_ci_male$percent[4],
        post_ci_male$percent[4],        
        delta_ci_male$percent[4]
        ),
    hi = c(
        pre_ci_fempre$percent[5],
        post_ci_fempre$percent[5],        
        delta_ci_fempre$percent[5],
        pre_ci_fempost$percent[5],
        post_ci_fempost$percent[5],        
        delta_ci_fempost$percent[5],
        pre_ci_male$percent[5],
        post_ci_male$percent[5],        
        delta_ci_male$percent[5]
        )
)

INRmedians$lo <- round(INRmedians$lo, 2)
INRmedians$hi <- round(INRmedians$hi, 2)

premd <- INRmedians[c(1, 4, 7), ]
postmd <- INRmedians[c(2, 5, 8), ]
deltamd <- INRmedians[c(3, 6, 9), ]

# 0.99 quantile for x-limits
max_q975 <- INRs %>%
    group_by(sex_group_patient, variable) %>%
    summarise(q975 = quantile(INR, p=0.975)) %>%
    pull(q975) %>%
    max()
# Lower limit
min_INR <- INRs %>%
    group_by(sex_group_patient, variable) %>%
    summarise(min = min(INR, na.rm=TRUE)) %>%
    pull(min) %>%
    min()

# Ns for plot
INRs_N <- INRs %>%
    filter(INR <= max_q975) %>%
    group_by(sex_group_patient) %>%
    mutate(N = round(n() / 2)) %>%
    filter(variable == "pre") %>%
    select(sex_group_patient, variable, N) %>%
    distinct()

# df for "Medians [99% CI]" title lol
titles <- data.frame(
    lab = rep("Medians [99% CI]", 3),
    sex_group_patient = factor(c("Female | Premenopausal", "Female | Postmenopausal", "Male"), levels = c("Female | Premenopausal", "Female | Postmenopausal", "Male")),
    variable = rep("pre", 3)
    )

ggplot(INRs, aes(x = INR, fill = factor(variable, levels = c("pre", "post")))) +
    geom_density() +
    scale_fill_manual(values = gray(c(0.7, 0.3), 0.5), name = "", labels = c("Before transfusion", "After transfusion")) +
    geom_text(data=titles, aes(label = "Medians [99% CI]", x = 2, y = 1.75), fontface = "bold", size = 3, hjust = 0) +
    geom_text(data = INRs_N, aes(label = paste("N:", format(N, big.mark = ",", trim = T)), fontface = "bold", x = 1.5, y = 1.4), size = 4, hjust = 0) +
    geom_text(data = premd, 
              aes(label = paste0(
                    "Before transfusion(s): ", md, " [", lo,", ", hi, "]"
                    ), 
                  x = 2, 
                  y = 1.4
                 ), 
              size = 3, 
              hjust = 0
             ) +
    geom_text(data = postmd, 
            aes(label = paste0(
                "After transfusion(s): ", md, " [", lo,", ", hi, "]"
                ), 
                x = 2, 
                y = 1.05
                ), 
            size = 3, 
            hjust = 0
            ) +
    geom_text(data = deltamd, 
            aes(label = paste0(
                "Δ per transfused unit: ", round(md, 4), " [", lo,", ", hi, "]"
                ), 
                x = 2, 
                y = 0.7
                ), 
            size = 3, 
            hjust = 0
            ) +
    geom_vline(data = premd, aes(xintercept = md), linetype = "dotdash") +
    geom_vline(data = postmd, aes(xintercept = md), linetype = "dashed") +
    facet_wrap(vars(sex_group_patient), ncol = 1) +
    theme_minimal() +
    labs(y = "Density", x="INR") +
    scale_x_continuous(breaks = seq(0.5, max_q975, by=0.5), limits = c(min_INR, max_q975)) +
    theme(legend.position = "bottom", strip.text = element_text(size = 13), axis.text = element_text(size = 13), axis.title = element_text(size = 13), legend.text = element_text(size = 13), axis.title.x = element_text(margin = margin(t = 14)))

ggsave("./results/manuscript1_plots/S5_B.png", width = 2000, height = 2000, units = "px", bg = "white")
ggsave("./results/manuscript1_plots/svg/S5_B.svg", width = 2000, height = 2000, units = "px", bg = "white")

# Results computed over
100 * sum(INRs$epis_n) / 2 / nrow(master) #%
# of units
```

## Blood types by product

```{r}
master <- master %>%
    mutate(
        bloodType_ID = paste0(abo_patient, " ", rhd_patient),
        bloodType_unit = paste0(abo_product, " ", rhd_product)
    )
master$abo_patient <- factor(master$abo_patient, levels = c("O", "A", "B", "AB"))
master$abo_product <- factor(master$abo_product, levels = c("O", "A", "B", "AB"))
master$rhd_patient <- factor(master$rhd_patient, levels = c("pos", "neg"))
master$rhd_product <- factor(master$rhd_product, levels = c("pos", "neg"))

master$bloodType_ID <- factor(master$bloodType_ID, levels = c("AB pos", "AB neg", "A pos", "A neg", "B pos", "B neg", "O pos", "O neg", "A NA", "B NA", "O NA", "NA pos", "NA neg", "NA NA"))
master$bloodType_unit <- factor(master$bloodType_unit, levels = c("AB pos", "AB neg", "A pos", "A neg", "B pos", "B neg", "O pos", "O neg", "A NA", "B NA", "O NA", "NA pos", "NA neg", "NA NA"))

master$type_product <- factor(master$type_product, levels = c("RBC", "Platelets", "Plasma"))

master %>%
    filter(type_product == "RBC") %>%
    filter( !is.na(abo_patient) & !is.na(abo_product) & !is.na(rhd_patient) & !is.na(rhd_product)) %>%
    ggplot(aes(x = bloodType_ID, fill = bloodType_unit)) +
        geom_bar() +
        facet_wrap(~type_product, ncol=1)

abo_plot <- master %>%
    filter(!is.na(type_product)) %>%
    filter(!is.na(abo_patient) & !is.na(abo_product)) %>%
    group_by(type_product, abo_patient, abo_product) %>%
    filter(n_distinct(id_patient) >= 5) %>%
    summarise(n = n())

bg_colors = c("#00BFB2", "#e7298a", "#e6ab02", "#342e37")

ggplot(abo_plot, aes(x = abo_patient, y = n/sum(abo_plot$n)*100, fill = abo_product)) +
        geom_bar(stat="identity") +
        facet_wrap(~type_product, nrow=1) +
        scale_fill_manual(values = bg_colors, name = "ABO group of\ntransfused unit") +
        scale_y_continuous(expand = c(0.01,0.01), breaks = seq(0,25,by=5)) +
        theme_minimal() +
        theme(
            axis.text.x = element_text(size = 12),
            axis.text.y = element_text(size = 12),
            text = element_text(size = 12),
            legend.position="top", 
            axis.title.x = element_text(margin = margin(t = 14)),
            strip.text = element_text(size = 12)
        ) +
        labs(x = "ABO group of patient", y = "Percent of transfused units (%)")


ggsave("./results/manuscript1_plots/S6.png", width = 3000, height = 1500, bg = "white", units = "px")
ggsave("./results/manuscript1_plots/svg/S6.svg", width = 3000, height = 1500, bg = "white", units = "px")
```
