---
title: "Explore new DRG files"
format: html
---

```{r}
# Libraries
library(reticulate) # for pyenv stuff
pandas <- import("pandas") # for fastparquet to work
library(parallel) # for detectCores
library(doParallel) # for registerDoParallel
library(tidyr) # for unnest
library(dplyr)
library(ggplot2) # for plotting
library(purrr) # for map_int
```

# Load data

Not all will probably fit into memory simultaneously.

```{r}
og <- readRDS("./results/data/jakso_drg_koodi.rds")
names(og)
min(og$pvmalku)
max(og$pvmalku)
```

```{r}
new1 <- readRDS("./results/data/drg2024_pre2022_1.rds")
names(new1) # different columns than in og
min(new1$pvmalku)
max(new1$pvmalku)

new2 <- readRDS("./results/data/drg2024_pre2022_2.rds")
names(new2) # different columns than in new1
min(new2$alkuhetki)
max(new2$alkuhetki)
```

```{r}
new3 <- readRDS("./results/data/drg2024_post2022.rds")
names(new3) # again different columns
min(new3$alkuhetki, na.rm = T)
max(new3$alkuhetki, na.rm = T) # 2024-05-22!! >:(
```

```{r}
ggplot(new3, aes(x = alkuhetki)) +
    geom_density(alpha = 0.6) +
    theme_minimal()

ggsave("newDRG_timedist.png", bg = "white")
```

# Combined

## Select relevant columns

NB! All three of these contain all sorts of useful information (like diag/eala) that gets dropped here. We might return to look at it at a later point, but currently we're interested in the actual DRGs, SHJAKSO_NUM and time stamps.

drg:
new1 -- `drg_alkup`
new2 -- `drg`
new3 -- `Drg_alkup`

time stamps:
new1 -- `pvmalku` and `pvmlahto`
new2 -- `alkuhetki` and `loppuhetki`
new3 -- `alkuhetki` and `loppuhetki`

shjakso_numero:
new1 -- `id_oshoito_kaynti`
new2 -- `shjakso_numero` (NB!! Also has `id_oshoito_kaynti` but they don't match to shnum elsewhere, unlike in new1)
new3 -- `shjakso_numero`

```{r}
new1_sel <- new1 %>%
    select(henkilotunnus, drg_alkup, pvmalku, pvmlahto, id_oshoito_kaynti) %>%
    rename(drg = drg_alkup, alkuhetki = pvmalku, loppuhetki = pvmlahto, shjakso_numero = id_oshoito_kaynti)
new2_sel <- new2 %>%
    select(henkilotunnus, drg, alkuhetki, loppuhetki, shjakso_numero)
new3_sel <- new3 %>%
    select(henkilotunnus, Drg_alkup, alkuhetki, loppuhetki, shjakso_numero) %>%
    rename(drg = Drg_alkup)
rm(new1, new2, new3)
```

```{r}
new_drgs <- bind_rows(new1_sel, new2_sel, new3_sel)

saveRDS(new_drgs, "./results/data/new_drgs_neat.rds")
```